---
title: "Getting started with DRUID"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with DRUID}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Overview

DRUID scores drugs against a differential expression signature using:

1. A **corpus** of drug transcriptional profiles (gene-direction tokens)
2. **TF-IDF weighting** (default: geometric mean of dual views)
3. **Cosine similarity** between query and each drug profile
4. An **empirical null** and composite **DRUID score**

## Minimal example (bundled CMAP)

```{r}
library(DRUID)

# Synthetic illustration — replace with real limma/DESeq2 output
entrez <- c("7157", "1956", "5290", "207")
dge_matrix <- cbind(
  c(2.1, -1.5, 1.8, -2.3),   # log2FC
  c(0.001, 0.01, 0.002, 0.005) # p-value
)

res <- concoct(
  dge_matrix = dge_matrix,
  entrez = entrez,
  dataset = "cmap",
  tfidf_mode = "geom_tf",
  druid_direction = "neg",
  fold_thr = 0,
  pvalue_thr = 0.05,
  num_random = 500
)

head(res[, c("drug_name", "cell_line", "number_matches", "druid_score")])
```

## Corpus modes

- `geom_tf` — recommended; geometric mean of drug- and gene-as-document TF-IDF
- `combined` — historical Hadamard product
- `binary` — unweighted overlap
- `drug_tfidf` — single-view drug TF-IDF

## Performance

For large `num_random` (e.g. 10,000), use parallel null evaluation:

```{r}
res <- concoct(..., num_random = 10000, n_cores = 4)
```

## Further reading

- `README.md` — API reference
- `inst/extdata/DATA_SOURCES.md` — data provenance and versioning
- `ablation_results/ablation_report.md` — corpus mode comparison
